Clustering by category and plots

### 2024-10-04 data/04-bycategoriy-clustering.qmd
Updates including Lucias comments

# Not caring - not complete (C2)
# Somewhat caring - not complete (C1)
# Somewhat caring - complete (C4)
# Caring - somewhat complete (C3)
# Caring - very complete (C6)
# Very caring - complete (C5)
# Very caring - very complete (C7)
```
```{r}
#par(mfrow=c(3,2))
plot(access.som_small, 
     what = "obs", 
     type = "color", 
     variable = 14, 
     show.names=TRUE, 
     main="norm_Dependent_centric")
plot(access.som_small, 
     what = "obs", 
     type = "color", 
     variable = 17, 
     show.names = TRUE, 
     main = "norm_Errand_centric")
# plot(access.som_small, 
#      what = "obs", 
#      type = "color", 
#      variable = 3,
#      show.names = TRUE, 
#      main="norm_Grocery_centric")
# plot(access.som_small, 
#      what="obs", 
#      type="color", 
#      variable=4, 
#      show.names = TRUE, 
#      main="norm_Health_centric")
# plot(access.som_small, 
#      what="obs", 
#      type="color",
#      variable = 9, 
#      show.names = TRUE, 
#      main="Diversity")

@@

```{r}
access_15_wide |> select(total_tot_copp_isc_groc,total_tot_copp_isc_dep,total_tot_copp_isc_health,total_tot_copp_isc_err,div_allcats) |> summary()
access_15_wide |> group_by(scluster) |> summarise(total_tot_copp_isc_groc = median(total_tot_copp_isc_groc),
                                                  total_tot_copp_isc_dep = median(total_tot_copp_isc_dep),
                                                  total_tot_copp_isc_health = median(total_tot_copp_isc_health),
                                                  total_tot_copp_isc_err = median(total_tot_copp_isc_err),
                                                  div_allcats = median(div_allcats),
                                                  pc = n()/143882)
```

@@

The low diversities - low accessibility
C7 - the WORST - below Q3 in everything

Mixed diversities - Higher accessibility for some stuff
C4 - above median dependent-centric - but low diversity and low everything else (THESE AREAS HAVE POTENTIAL! the need more GROC, HEALTH, and ERRs) 
C5 - like medium - about median everything with slightly lower HIGH diversity
C3 - exceptionally high dependent-centric, and high everything else but low diversity

High diversities and high accessibility
C6 - medium - about median everything with the highest diversity
C2 - better than medium - slightly above median everything with high diversity
C1 - the best - above Q3 everything

THE LABELS:
Not caring - not complete (C7)
Somewhat caring - not complete (C4)
Somewhat caring - complete (C5)
Caring - somewhat complete (C3)
Caring - very complete (C6)
Very caring - complete (C2)
Very caring - very complete (C1)

@@

df <- df |> 
  mutate(Total_copp_groc = Convience +Grocer,
         Total_copp_dep = Rec + Daycare + LTC + Park + School + Senior,
         Total_copp_health = Hos_clinc+Dentist,
         Total_copp_err = PostOff +Library+Bank_ATM + Pharmacy)
```

Train a decision tree using the input SOM input variables:
```{r}
# classmod.som <- rpart(data = df,
#                       formula = scluster ~ (Convience + Grocer + div_groc +
#                                               Rec + Daycare + LTC + Park + School + Senior + div_dep +
#                                               Dentist + Hos_clinc + div_health +
#                                               PostOff +Library+Bank_ATM + Pharmacy + div_err))
classmod.som <- rpart(data = df,
                      formula = scluster ~ (norm_tot_copp_isc_norm_groc + norm_Convience + norm_Grocer + #div_allcats +
                                              norm_tot_copp_isc_norm_dep + norm_Rec + norm_Daycare + norm_LTC + norm_Park + norm_School + norm_Senior + #div_dep +
                                              norm_tot_copp_isc_norm_health + norm_Pharmacy + norm_Dentist + norm_Hos_clinc + #div_health +
                                              norm_tot_copp_isc_norm_err + norm_PostOff + norm_Library+ norm_Bank_ATM + div_allcats#+ div_err
                                            ))

@@

df$SC1 <- predicted_labels$`1`
df$SC2 <- predicted_labels$`2`
df$SC3 <- predicted_labels$`3`
df$SC4 <- predicted_labels$`4`
df$SC5 <- predicted_labels$`5`
df$SC6 <- predicted_labels$`6`
df$SC7 <- predicted_labels$`7`
# df$SC8 <- predicted_labels$`8`
# df$SC9 <- predicted_labels$`9`

@@

Trying ggparty to visualise, but first convert the rpart object to a party object:
```{r}
classmod.sompar <- as.party(classmod.som)

#Autoplot as ggplot:
autoplot(classmod.sompar)
```

Frequency of parcels in the super clusters:
```{r}
freq_sc <- ggplot(access_15_wide |> 
         mutate(scluster = factor(scluster))) +
  geom_bar(aes(x = "", 
               group = scluster, 
               fill = scluster)) +
  scale_fill_discrete(name = "Supercluster") #+
  #theme_void() +
  #guides(fill = guide_legend(nrow = 1)) +
  #theme(legend.position = "bottom")

#with new names and colours
freq_sc_renamed <- ggplot(access_15_wide |> 
         mutate(scluster = factor(scluster))) +
  geom_bar(aes(x = "", 
               group = scluster, 
               fill = scluster)) +
  scale_fill_manual(name = "Neighbourhood typologies",
                    values = c( "#1A9850","#91CF60", "#D9EF8B", "#FFFFBF","#FEE08B","#FC8D59","#D73027"),
                    labels = c("Not caring - not complete (C2)",
                               "Somewhat caring - not complete (C1)",
                               "Somewhat caring - complete (C4)",
                               "Caring - somewhat complete (C3)",
                               "Caring - very complete (C6)",
                               "Very caring - complete (C5)",
                               "Very caring - very complete (C7)"))

Extract the legend:
```{r}
legd <- cowplot::get_legend(freq_sc)

legd
```

Make a prettier plot:
```{r}
party_plot <- ggparty(classmod.sompar) +
  geom_edge() +
  geom_edge_label(size = 1.6) +
  geom_node_label(aes(label = splitvar),
                  id = "inner", 
                  size = 2, nudge_y = -0.022) +
  # pass list to gglist containing all ggplot components we want to plot for each
  # (default: terminal) node
  geom_node_plot(gglist = list(geom_bar(aes(x = "", fill = scluster),
                                        position = position_fill()),
                               xlab(""),
                               ylab(""),
                               theme_void()
                               #theme(axis.text = element_blank())
                               ))
party_plot
```

Plot the tree:
```{r}
png(file= file.path(here("figures/decision_tree_profiles_sc.png")),width= 2404, height= 1600, units="px",  res=300)
plot_grid(party_plot, legd, nrow = 1, rel_widths = c(1, 0.2))
dev.off()
```

@@

```{r}
df_DA <- df |> left_join(access_15 |> dplyr::select(c("from_id","GeoUID")), by = "from_id") |> 
  group_by(GeoUID) |> 
  summarise(
    SC1 = median(SC1),
    SC2 = median(SC2),
    SC3 = median(SC3),
    SC4 = median(SC4),
    SC5 = median(SC5),
    SC6 = median(SC6),
    SC7 = median(SC7),
    parcel_count = n_distinct(from_id),
    max_label_val = pmax(SC1, SC2, SC3, SC4, SC5, SC6, SC7),
    div_allcats = median(div_allcats),
    Total_copp_groc = median(Total_copp_groc),
    Total_copp_dep = median(Total_copp_dep),
    Total_copp_health = median(Total_copp_health),
    Total_copp_err = median(Total_copp_err)) |>
  mutate(max_label = case_when(max_label_val == SC1 ~ "a. Very caring - very complete (SC1)",
                               max_label_val == SC2 ~ "b. Very caring - mostly complete (SC2)",
                               max_label_val == SC3 ~ "d. Somewhat caring - mostly complete (SC3)",
                               max_label_val == SC4 ~ "c. Very caring - mostly complete (SC4)",
                               max_label_val == SC5 ~ "j. not caring - somewhat complete (SC5)",
                               max_label_val == SC6 ~ "e. Somewhat caring - mostly complete (SC6)",
                               max_label_val == SC7 ~ "f. not caring - not complete (SC7)"))

                    labels = c("Not caring - not complete (C2)",
                               "Somewhat caring - not complete (C1)",
                               "Somewhat caring - complete (C4)",
                               "Caring - somewhat complete (C3)",
                               "Caring - very complete (C6)",
                               "Very caring - complete (C5)",
                               "Very caring - very complete (C7)"))

@@

df_DA$parcel_count |> sum() #a check
```
Loading boundaries for mapping:
```{r}
Community_Boundaries <- st_read("data-raw/boundaries/Community_Boundaries.shp") |> mutate(COMMUNITY_ = ifelse(COMMUNITY_ == "Hamilton", "Hamilton-Central", COMMUNITY_))|> st_transform(crs=4326)
City_Boundary <- st_read("data-raw/boundaries/City_Boundary.shp") |> st_transform(crs=4326)
hydro_p_LakeOntario <- st_read("data-raw/boundaries/hydro_p_LakeOntario.shp") |> st_transform(crs=4326)
ham_bay <- st_read("data-raw/boundaries/Waterbodies.shp") |> st_transform(crs=4326) |> filter(FEATURE_TY == "Lake")
ham_bay_cropped<-st_crop(ham_bay$geometry, Community_Boundaries$geometry)
hydro_p_LakeOntario_cropped<-st_crop(hydro_p_LakeOntario$geometry, Community_Boundaries$geometry)
```

@@

```{r, eval=FALSE}
total_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(df_DA) +
  tm_polygons("Total_copp_groc",
              border.alpha = 0.2,
              #style = "quantile", n=4,
              palette = "Purples",
              title = "Total access to care destinations",) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4,
            bg.color = "grey",
            legend.position = c("right","top"))

total_copp_plot
```

```{r, eval=FALSE}
SClabel_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(df_DA) +
  tm_polygons("max_label",
              border.alpha = 0.2,
              # palette = c("#44ce1b","#bbdb44", "#f7e379", "#f2a134","#e51f1f"),
              # labels = (c("Very caring - very complete",
              #             "Caring - Complete",
              # "Somewhat caring - complete",
              # "Somewhat caring - somewhat complete", 
              # "Somewhat caring - somewhat complete")),
              title = "Neighbourhood typology") +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4,
            bg.color = "grey",
            legend.position = c("right","top"))
SClabel_copp_plot

SCLab_and_total_copp_plot <- tmap_arrange(SClabel_copp_plot, total_copp_plot, nrow=2)
```

```{r, eval=FALSE}
SCLab_and_total_copp_plot
```

```{r, eval=FALSE}
tmap_save(SCLab_and_total_copp_plot, 
          file= file.path(here("figures/SCLab_and_total_copp_plot.png")),
          dpi=300)

tmap_save(SClabel_copp_plot, 
          file= file.path(here("figures/SClabel_copp_plot.png")),
          dpi=300)
```

@@

Select data for classification:
```{r}

df <- access_15_wide |> left_join(access_15_aggregated, by = "from_id") |> 
  transmute(scluster = factor(scluster),
            Income_bottom = `v_CA21_1103: In bottom half of the distribution`/(`v_CA21_1103: In bottom half of the distribution` + `v_CA21_1121: In top half of the distribution`),
            LICO = `v_CA21_1085: Prevalence of low income based on the Low-income cut-offs, after tax (LICO-AT) (%)`,
            Avg_numb_of_kids = `v_CA21_498: Average number of children in census families with children`,
            Not_in_labforce = `v_CA21_6504: Not in the labour force`/(`v_CA21_6498: Employed` + `v_CA21_6501: Unemployed` + `v_CA21_6504: Not in the labour force`),
            Income_gini = `v_CA21_1142: Gini index on adjusted household after-tax income`,
            Vis_minority = `v_CA21_4875: Total visible minority population`/ (`v_CA21_4914: Not a visible minority`+ `v_CA21_4875: Total visible minority population`),
            Age_0to14 =(`v_CA21_11: 0 to 14 years`) / (`v_CA21_11: 0 to 14 years` + `v_CA21_68: 15 to 64 years` + `v_CA21_251: 65 years and over`),
            One_parent_families = `v_CA21_507: Total one-parent families` / (`v_CA21_507: Total one-parent families` + `v_CA21_500: Total couple families`),
            walk_towork =  `v_CA21_7647: Walked` / (`v_CA21_7647: Walked` + `v_CA21_7635: Car, truck or van` + `v_CA21_7644: Public transit` + `v_CA21_7650: Bicycle` + `v_CA21_7653: Other method` ),
            Owner_hld_CHN = `v_CA21_4308: % in core housing need (57)`,
            Hld_ownership = `v_CA21_4306: % of owner households with a mortgage (58)`,
            Tenant_subsi_hld = `v_CA21_4314: % of tenant households in subsidized housing (61)`,
            Tenant_hld_CHH =`v_CA21_4316: % in core housing need (57)`,
            Population,
            Low_or_no_deg = (`v_CA21_5868: No certificate, diploma or degree` + `v_CA21_5871: High (secondary) school diploma or equivalency certificate`) / (`v_CA21_5868: No certificate, diploma or degree` + `v_CA21_5871: High (secondary) school diploma or equivalency certificate` + `v_CA21_5874: Postsecondary certificate, diploma or degree`))

# examining correlations
cor(df |> select(-c("scluster")), use = "complete.obs")

#Train a decision tree using the data (renamed!):
classdemo.som <- rpart(data = df,
                      formula = scluster ~ (Income_bottom + Age_0to14 + One_parent_families + Vis_minority + Not_in_labforce+ Income_gini+walk_towork+Owner_hld_CHN+Hld_ownership+Tenant_subsi_hld+Tenant_hld_CHH+LICO),
                      weights = Population
                      )
# # Visualize the decision tree:
# rpart.plot(classdemo.som, cex = 0.5, type = 0, varlen = 8)
```

>>>>

# Not caring - not complete (C2)
# Somewhat caring - not complete (C1)
# Somewhat caring - complete (C4)
# Caring - somewhat complete (C3)
# Caring - very complete (C6)
# Very caring - complete (C5)
# Very caring - very complete (C7)
```

```{r}
#par(mfrow=c(3,2))
plot(access.som_small, 
     what = "obs", 
     type = "color", 
     variable = 14, 
     show.names=TRUE, 
     main="norm_Dependent_centric")
plot(access.som_small, 
     what = "obs", 
     type = "color", 
     variable = 17, 
     show.names = TRUE, 
     main = "norm_Errand_centric")
# plot(access.som_small, 
#      what = "obs", 
#      type = "color", 
#      variable = 3,
#      show.names = TRUE, 
#      main="norm_Grocery_centric")
# plot(access.som_small, 
#      what="obs", 
#      type="color", 
#      variable=4, 
#      show.names = TRUE, 
#      main="norm_Health_centric")
# plot(access.som_small, 
#      what="obs", 
#      type="color",
#      variable = 9, 
#      show.names = TRUE, 
#      main="Diversity")

@@

```{r}
# norm_tot_copp_isc_norm_groc + norm_Convience + norm_Grocer + #div_allcats +
#                                               norm_tot_copp_isc_norm_dep + norm_Rec + norm_Daycare + norm_LTC + norm_Park + norm_School + norm_Senior + #div_dep +
#                                               norm_tot_copp_isc_norm_health + norm_Pharmacy + norm_Dentist + norm_Hos_clinc + #div_health +
#                                               norm_tot_copp_isc_norm_err + norm_PostOff + norm_Library+ norm_Bank_ATM + div_allcats

#total_tot_copp_isc_dep,total_tot_copp_isc_health,total_tot_copp_isc_err,div_allcats
#'total_Community or Recreation Centre', "total_Daycare or EarlyON","total_Long_Term Care or Retirement Home","total_Park","total_School",'total_Senior Centre'

access_15_wide |> 
  select(c(
    scluster,total_tot_copp_isc_groc,total_tot_copp_isc_dep,total_tot_copp_isc_health,total_tot_copp_isc_err,div_allcats
    #scluster,norm_tot_copp_isc_norm_groc, norm_tot_copp_isc_norm_dep,norm_tot_copp_isc_norm_health,norm_tot_copp_isc_norm_err,div_allcats
    )) |> 
  mutate(tot = rowSums(across(c(total_tot_copp_isc_dep,total_tot_copp_isc_health,total_tot_copp_isc_err))))|>
  group_by(scluster) |>
  summarise_all(list(median))|>
    mutate(lab = case_when(scluster == 1 ~ "A+",
                               scluster == 2 ~ "A",
                               scluster == 3 ~ "A-",
                               scluster == 4 ~ "D",
                               scluster == 5 ~ "B",
                               scluster == 6 ~ "B+",
                               scluster == 7 ~ "F")) |> arrange(tot)

  # total_tot_copp_isc_groc = mean(total_tot_copp_isc_groc),
  #                                                 total_tot_copp_isc_dep = mean(total_tot_copp_isc_dep),
  #                                                 total_tot_copp_isc_health = mean(total_tot_copp_isc_health),
  #                                                 total_tot_copp_isc_err = mean(total_tot_copp_isc_err),
  #                                                 div_allcats = mean(div_allcats),
                                                  #pc = n()/143882) |>
# access_15_wide |> 
#   select(c(scluster,norm_tot_copp_isc_norm_groc, norm_tot_copp_isc_norm_dep,norm_tot_copp_isc_norm_health,norm_tot_copp_isc_norm_err,div_allcats)) |> summary()
```

@@

The low diversities - low accessibility

Lower accessibility (under median) with mixed diversity
F - C7 - the absolute worst, low in all but with some dep, making it very low diversity - "NOT GOOD - need to add it all"
D - C4 - has nothing except dep, making it very low diversity - "NOT GOOD - need to add err, groc, health"

B - C5 - LIKE C6 but slightly lower everything. Overall good balance so good diversity - "ALL AROUND GOOD" -- firmly average. 
B+ - C6 - about median dep, above median groc and health and Q3 errand. Overall great balance so high diversity - "ALL AROUND GREAT"

High accessibilities (high caring) -- all at or above median
A- - C3 - about median groc, health, errand and Q3 dep-centric, so lower diversity score - "ALL AROUND GREAT - with really high children-centric dests" 
A - C2 - above Q3 everything and high diversity - however all is slightly lower than A - "ALL AROUND EXCELLENT"
A+ - C1 - the best - above Q3 everything with high diversity- "ALL AROUND SUPER EXCELLENT"

THE LABELS:

F (Nothing much: Not caring & not complete) (C7)
D (Daycares, parks and schools, not much else: Somewhat caring & not complete) (C4)

B (Average: Caring & complete) (C5)
B+ (Above average: Caring & very complete) (C6)

A- (Excellent, but dep- heavy: Very caring & not complete (C3)
A (Excellent: Very caring & very complete) (C2)
A+ (Exceptional: very caring & very complete) (C1)

@@

df <- df |> 
  mutate(Total_copp_groc = Convience +Grocer,
         Total_copp_dep = Rec + Daycare + LTC + Park + School + Senior,
         Total_copp_health = Hos_clinc+Dentist,
         Total_copp_err = PostOff +Library+Bank_ATM + Pharmacy) |>
  rename("GROCERY TOTALS" = "total_tot_copp_isc_groc",
             "DEPENDENT TOTALS"= total_tot_copp_isc_dep,
             "HEALTH TOTALS"= total_tot_copp_isc_health,
             "ERRAND TOTALS"= total_tot_copp_isc_err, 
             "CATEGORY DIVERSITY"=div_allcats)
```

Train a decision tree using the input SOM input variables:
```{r}
# classmod.som <- rpart(data = df,
#                       formula = scluster ~ (Convience + Grocer + div_groc +
#                                               Rec + Daycare + LTC + Park + School + Senior + div_dep +
#                                               Dentist + Hos_clinc + div_health +
#                                               PostOff +Library+Bank_ATM + Pharmacy + div_err))

classmod.som <- rpart(data = df,
                      formula = scluster ~ (`GROCERY TOTALS` +
                                            #norm_tot_copp_isc_norm_groc + 
                                              #norm_Convience + norm_Grocer + #div_allcats +
                                              `DEPENDENT TOTALS` +
                                              #norm_tot_copp_isc_norm_dep + 
                                              #norm_Rec + norm_Daycare + norm_LTC + norm_Park + norm_School + norm_Senior + #div_dep +
                                              `HEALTH TOTALS` +
                                              #norm_tot_copp_isc_norm_health + 
                                              #norm_Pharmacy + norm_Dentist + norm_Hos_clinc + #div_health +
                                              `ERRAND TOTALS` + 
                                              #norm_tot_copp_isc_norm_err + 
                                              #norm_PostOff + norm_Library+ norm_Bank_ATM + 
                                              `CATEGORY DIVERSITY`#+ div_err
                                            ))

@@

df$SC1 <- predicted_labels$`1`
df$SC2 <- predicted_labels$`2`
df$SC3 <- predicted_labels$`3`
df$SC4 <- predicted_labels$`4`
df$SC5 <- predicted_labels$`5`
df$SC6 <- predicted_labels$`6`
df$SC7 <- predicted_labels$`7`
# df$SC8 <- predicted_labels$`8`
# df$SC9 <- predicted_labels$`9`
```

Variable importance:
```{r}
var_importance <- data.frame(imp = classmod.som$variable.importance)
var_importance <- var_importance %>% 
  tibble::rownames_to_column() %>% 
  dplyr::rename("variable" = rowname) %>% 
  dplyr::arrange(imp) %>%
  dplyr::mutate(variable = forcats::fct_inorder(variable))
ggplot2::ggplot(var_importance) +
  geom_col(aes(x = variable, y = imp),
           col = "black", show.legend = F) +
  coord_flip() +
  scale_fill_grey() +
  theme_bw()

@@

<!-- Trying ggparty to visualise, but first convert the rpart object to a party object: -->
<!-- ```{r} -->
<!-- classmod.sompar <- as.party(classmod.som) -->

<!-- #Autoplot as ggplot: -->
<!-- autoplot(classmod.sompar) -->
<!-- ``` -->

<!-- Frequency of parcels in the super clusters: -->
<!-- ```{r} -->
<!-- freq_sc <- ggplot(access_15_wide |>  -->
<!--          mutate(scluster = factor(scluster))) + -->
<!--   geom_bar(aes(x = "",  -->
<!--                group = scluster,  -->
<!--                fill = scluster)) + -->
<!--   scale_fill_discrete(name = "Supercluster") #+ -->
<!--   #theme_void() + -->
<!--   #guides(fill = guide_legend(nrow = 1)) + -->
<!--   #theme(legend.position = "bottom") -->

<!-- #with new names and colours -->
<!-- freq_sc_renamed <- ggplot(access_15_wide |>  -->
<!--          mutate(scluster = factor(scluster))) + -->
<!--   geom_bar(aes(x = "",  -->
<!--                group = scluster,  -->
<!--                fill = scluster)) + -->
<!--   scale_fill_manual(name = "Neighbourhood typologies", -->
<!--                     values = c( "#1A9850","#91CF60", "#D9EF8B", "#FFFFBF","#FEE08B","#FC8D59","#D73027"), -->
<!--                     labels = c("F (Nothing much: Not caring & not complete) (C7)", -->
<!-- "D (Daycares, parks and schools, not much else: Somewhat caring & not complete) (C4)", -->
<!-- "B (Average: Caring & complete) (C5)", -->
<!-- "B+ (Above average: Caring & very complete) (C6)", -->
<!-- "A- (Excellent, but dep- heavy: Very caring & not complete (C3)", -->
<!-- "A (Excellent: Very caring & very complete) (C2)", -->
<!-- "A+ (Exceptional: very caring & very complete) (C1)")) -->

<!-- ``` -->

<!-- Extract the legend: -->
<!-- ```{r} -->
<!-- legd <- cowplot::get_legend(freq_sc) -->

<!-- legd -->
<!-- ``` -->

Make a prettier plot:
```{r}
classmod.sompar <- as.party(classmod.som)

party_plot <- ggparty(classmod.sompar) +
  geom_edge() +
  geom_edge_label(size = 2.5) +
  geom_node_label(aes(label = splitvar ),
                  id = "inner",
                  size = 2, nudge_y = -0.022) +
  # pass list to gglist containing all ggplot components we want to plot for each
  # (default: terminal) node
  geom_node_plot(gglist = list(geom_bar(aes(x = "", fill = scluster),
                                        position = position_fill()),
                               xlab(""),
                               ylab(""),
                               theme_void(),
                               scale_fill_manual(values = c("#1A9850","#91CF60", "#D9EF8B", "#F46D43", "#FDAE61", "#FEE08B","#D73027")),
                              theme(axis.text = element_blank())))
party_plot
```

Plot the tree:
```{r}
png(file= file.path(here("figures/access_sc_profiles_usingSCVARS.png")),width= 3000, height= 1600, units="px",  res=300)
plot_grid(party_plot, legd_renamed, nrow = 1, rel_widths = c(2, 1.5))
dev.off()
```

@@

"F (Nothing much: Not caring & not complete) (C7)",
"D (Daycares, parks and schools, not much else: Somewhat caring & not complete) (C4)",
"B (Average: Caring & complete) (C5)",
"B+ (Above average: Caring & very complete) (C6)",
"A- (Excellent, but dep- heavy: Very caring & not complete (C3)",
"A (Excellent: Very caring & very complete) (C2)",
"A+ (Exceptional: very caring & very complete) (C1)"

```{r}
df_DA <- df |> left_join(access_15 |> dplyr::select(c("from_id","GeoUID")), by = "from_id") |> 
  group_by(GeoUID) |> 
  summarise(
    SC1 = median(SC1),
    SC2 = median(SC2),
    SC3 = median(SC3),
    SC4 = median(SC4),
    SC5 = median(SC5),
    SC6 = median(SC6),
    SC7 = median(SC7),
    parcel_count = n_distinct(from_id),
    max_label_val = pmax(SC1, SC2, SC3, SC4, SC5, SC6, SC7),
    div_allcats = median(div_allcats),
    Total_copp_groc = median(Total_copp_groc),
    Total_copp_dep = median(Total_copp_dep),
    Total_copp_health = median(Total_copp_health),
    Total_copp_err = median(Total_copp_err)) |>
  mutate(max_label = case_when(max_label_val == SC1 ~ "a. (Exceptional: very caring & very complete) (C1)",
                               max_label_val == SC2 ~ "b. (Excellent: Very caring & very complete) (C2)",
                               max_label_val == SC3 ~ "c. (Excellent, but dep- heavy: Very caring & not complete (C3)",
                               max_label_val == SC4 ~ "f. (Daycares, parks and schools, not much else: Somewhat caring & not complete) (C4)",
                               max_label_val == SC5 ~ "e. (Average: Caring & complete) (C5)",
                               max_label_val == SC6 ~ "d. (Above average: Caring & very complete) (C6)",
                               max_label_val == SC7 ~ "g. (Nothing much: Not caring & not complete) (C7)"))

@@

df_DA$parcel_count |> sum() #a check

save(df_DA, file = "df_DA.RData")
```
Loading boundaries for mapping:
```{r}
Community_Boundaries <- st_read("data-raw/boundaries/Community_Boundaries.shp") |> mutate(COMMUNITY_ = ifelse(COMMUNITY_ == "Hamilton", "Hamilton-Central", COMMUNITY_))|> st_transform(crs=4326)
City_Boundary <- st_read("data-raw/boundaries/City_Boundary.shp") |> st_transform(crs=4326)
hydro_p_LakeOntario <- st_read("data-raw/boundaries/hydro_p_LakeOntario.shp") |> st_transform(crs=4326)
ham_bay <- st_read("data-raw/boundaries/Waterbodies.shp") |> st_transform(crs=4326) |> filter(FEATURE_TY == "Lake")
ham_bay_cropped<-st_crop(ham_bay$geometry, Community_Boundaries$geometry)
hydro_p_LakeOntario_cropped<-st_crop(hydro_p_LakeOntario$geometry, Community_Boundaries$geometry)

Escpartment_Boundary <- st_read("data-raw/boundaries/Escarpment.shp", quiet = TRUE) |> st_transform(Escpartment_Boundary, crs = 4326) |> st_cast("LINESTRING")
Escpartment_Boundary_HAM <- Escpartment_Boundary |> st_intersection(City_Boundary) 
```

@@

```{r, eval=FALSE}
# RColorBrewer::brewer.pal(n = 8, name = "RdYlGn")
# "#D73027" "#F46D43" "#FDAE61" "#FEE08B" "#D9EF8B" "#A6D96A" "#66BD63" "#1A9850"

SClabel_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(df_DA) +
  tm_polygons("max_label",
              border.alpha = 0.2,
              palette = c("#1A9850","#91CF60", "#D9EF8B", "#FEE08B","#FDAE61","#F46D43","#D73027"),
              labels = c("(A+) Exceptional: very caring & very complete", #C1
                         "(A ) Excellent: very caring & very complete ", #(C2)
                         "(A-) Dep- heavy, otherwise excellent: very caring & not complete", #(C3)
                         "(B+) Above average: caring & very complete", # (C6)
                         "(B ) Average: caring & complete", #(C5)
                         "(B- ) Dep- heavy, otherwise below average: somewhat caring & not complete", # (C4)
                         "(D ) Below average: not caring & not complete"), # (C7)
              title = "Neighbourhood typology") +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_shape(Escpartment_Boundary) + tm_lines(col = "grey30", lty = "dashed", lwd=1.5, legend.show=TRUE)+
  tm_scale_bar(position = c("left", "top"), breaks = c(0, 1, 5, 10)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4,
            bg.color = "grey",
            legend.text.size=0.8,
            legend.title.size = 1.2,
            legend.width = 2,
            legend.position = c("left","bottom"))
SClabel_plot

tmap_save(SClabel_plot, 
          file= file.path(here("figures/SClabel_plot.png")),
          dpi=300)
```

F (Nothing much: Not caring & not complete) (C7)
D (Daycares, parks and schools, not much else: Somewhat caring & not complete) (C4)

B (Average: Caring & complete) (C5)
B+ (Above average: Caring & very complete) (C6)

A- (Excellent, but dep- heavy: Very caring & not complete (C3)
A (Excellent: Very caring & very complete) (C2)
A+ (Exceptional: very caring & very complete) (C1)

@@

Select data for classification:

```{r}
# #min-max normalization
# rescale <- function(x, na.rm = TRUE) {
#   # Remove NA values if na.rm is TRUE
#   if (na.rm) {
#     min_x <- min(x, na.rm = TRUE)
#     max_x <- max(x, na.rm = TRUE)
#   } else {
#     min_x <- min(x)
#     max_x <- max(x)
#   }
#   
#   # Avoid division by zero (min == max) and ensure no Inf values
#   if (min_x == max_x) {
#     warning("All non-NA values are the same. Returning zeros.")
#     return(rep(0, length(x)))
#   }
#   
#   # Apply Min-Max normalization, keeping NAs where they were
#   rescaled_x <- (x - min_x) / (max_x - min_x)
#   
#   # Replace potential Inf/-Inf values with NA, though this should not happen
#   rescaled_x[is.infinite(rescaled_x)] <- NA
#   
#   return(rescaled_x)
# }
